---
title: "`r params$doc_title`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
params:
  doc_title: "Cholera Transmission Units: Model Results"
  downsample: F
  which_obs: "all_data"
  stan_chains: 4
  stan_iter_warmup: 250
  stan_iter_sample: 250
  run_model: T
  seed: 4
  save_plots: F
  min_year: 1970
  downsample_seed: 4
  sample_fraction: 0.2
  save_warmup: T
  sep_sudan: F
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, include = T, warning = F, root.dir = here::here())

```


## Stan HMM model on cholera transmission, real data:


```{r, echo = F}

run_model <- params$run_model
save_warmup <- params$save_warmup
stan_iter_sample <- params$stan_iter_sample
stan_iter_warmup <- params$stan_iter_warmup
stan_chains <- params$stan_chains

save_plots <- params$save_plots

seed <- params$seed
downsample <- params$downsample
downsample_seed <- params$downsample_seed
sample_fraction <- params$sample_fraction
sampling <- params$which_obs
min_year <- params$min_year
sep_sudan <- params$sep_sudan

simulate <- F
test_subset <- F

```

```{r initialize, message = F}

library(here)
library(cmdstanr)
library(bayesplot)
library(spData) 
library(sf)
library(posterior)
library(here)
library(cowplot)
library(tidyverse)

options(scipen = 999999999)

# Source functions
source(here("analysis","00_functions_settings.R"), local = T)

if (!dir.exists(out_dir)) {dir.create(out_dir)}

set.seed(seed)

```

Stan settings:

* Burn in: `r stan_iter_warmup` iterations
* Sampling: `r stan_iter_sample` iterations
* Chains: `r stan_chains`

<br>


```{r testing, include = F}
knitr::read_chunk(here("analysis","06_post_hmm_stats.R"))
```

```{r run_model, message = F, echo=F, include = T, warning = F}

```

```{r load_model, message = F, echo=F, include = F}

```

*Input: `r V` strains in `r M` countries over `r n_time` years.*

*Downsampling strategy: `r  paste(print_sample_func())`*


<br>

Summary diagnostics:

```{r , message = F}

model_objs$diagnostic_summary

```

<br>

```{r est_dfs, message = F, include = F, echo = F}
```

<br>

##### Rho: measure of probability of presence of each strain

<br>

```{r plot_rho, message = F, fig.height = 18, fig.width=12}
```

<br>

##### Prevalence

<br>

Inferred strain prevalence by country over time.

<br>

```{r plot_prev, message = F, fig.height = 18}
```


<br>

##### Gravity constant
$\kappa$ parameter

```{r}
model_fit_summary("kappa") |> select(-true)
traceplot(obj = model_objs$model_draws, "kappa")
mcmc_hist(model_objs$model_draws, pars = "kappa")
```

<br>

##### Spatial weights
$\omega_{i,j}$ parameters - asymmetric country-pair random effects

```{r map_prep, message = F, echo=F, include = F}
```

```{r map_w, message = F, echo=F}
```

<br>

##### Connectivity measure
$\xi_{i,j}$ parameters - asymmetric measure of between-country connectivity that is later used in clustering algorithms

```{r map_xi, message = F, echo=F}
```

<br>

##### Post-estimation inference

Cases attributed to each lineage using inferred lineage presence

```{r pred_cases, message = F, echo=F, fig.height=18}
```

<br>


##### Viterbi

<br>

Algorithm determining the most likely state sequence for each lineage by country and year.

<br>

```{r plot_zstar, message = F, fig.height = 18}
```

<br>
<br>

```{r pred_obs, message = F, echo = F, fig.width=14, fig.height = 15}
```

```{r save_clust, message = F, echo = F}
```


