---
title: "`r params$doc_title`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
params:
  doc_title: "Cholera Transmission Units: Model Diagnostics"
  downsample_random: F
  which_obs: "all_data"
  stan_chains: 4
  stan_iter_warmup: 250
  stan_iter_sample: 250
  seed: 4
  save_plots: F
  min_year: 1970
  downsample_seed: 4
  sample_fraction: 0.2
  save_warmup: F
  sep_sudan: F
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, include = T, warning = F, root.dir = here::here())

library(tidyverse)
library(cmdstanr)
library(spData) 
library(sf)
library(posterior)
library(here)
library(bayesplot)

```

```{r, include = F}

stan_iter_sample <- params$stan_iter_sample
stan_iter_warmup <- params$stan_iter_warmup
stan_chains <- params$stan_chains

save_plots <- params$save_plots
save_warmup <- params$save_warmup

seed <- params$seed
subsample <- params$subsample
downsample_random <- params$downsample_random
downsample_seed <- params$downsample_seed
sample_fraction <- params$sample_fraction
sampling <- params$which_obs
min_year <- params$min_year
sep_sudan <- params$sep_sudan

simulate <- F
test_subset <- F

```

```{r}

options(scipen = 999999999)

# Source functions
source(here("analysis","00_functions_settings.R"), local = T)

if (!dir.exists(out_dir)) {dir.create(out_dir)}

set.seed(seed)

```

## Diagnostics from HMM model:

<br>

Stan settings:

* Burn in: `r stan_iter_warmup` iterations
* Sampling: `r stan_iter_sample` iterations
* Chains: `r stan_chains`

<br>
```{r testing, include = F}
knitr::read_chunk(here("analysis","06_post_hmm_stats.R"))
```

```{r load_model, message = F, echo=F, include = F}

```

```{r , message = F, echo=F, include = F}

# parameter names
pars <- c("lp__", "eta", "kappa", "zeta", "tau[1]", "tau[2]", "delta") 

color_scheme_set("blue")

```

<br>

Run time (in hours):

<br>

```{r , message = F}

if(save_warmup){
model_objs$metadata$time |> as.data.frame() |> mutate_at(c("warmup", "sampling", "total"), ~ round(./(60*60),2))
}

```
<br>

Diagnostic summary:

<br>


```{r , message = F}

model_objs$diagnostic_summary

```

<br>

#### Pairs plots

<br>

```{r, message = F, fig.height = 10, fig.width = 10}

mcmc_pairs(model_objs$model_draws, pars = pars,
           off_diag_args = list(size = 0.5, alpha = 0.5),
           diag_fun = "dens")

```

<br>


#### Rhat

<br>

```{r, message = F, fig.height = 8}

mcmc_rhat(model_objs$rhats[pars[-1]]) + yaxis_text(hjust = 1)

```

<br>

#### Neff ratios

<br>

```{r, message = F, fig.height = 8}

mcmc_neff(model_objs$neffs[pars[-1]]) + yaxis_text(hjust = 1)

```

<br>

#### Parameter summaries:

<br>

##### Kappa (gravity constant)

```{r, message = F}

model_fit_summary("kappa") %>% select(-true)
traceplot(obj = model_objs$model_draws, "kappa")
mcmc_hist(model_objs$model_draws, pars = "kappa")

```

<br>

##### Zeta (distance transformation)

```{r, message = F}
model_fit_summary("zeta") %>% select(-true)
traceplot(obj = model_objs$model_draws, "zeta")
mcmc_hist(model_objs$model_draws, pars = "zeta")

```

<br>

##### Tau (population transformation)

```{r, message = F}

model_fit_summary("tau[1]") %>% select(-true)
model_fit_summary("tau[2]") %>% select(-true)
traceplot(obj = model_objs$model_draws, "tau")
mcmc_hist(model_objs$model_draws, regex_pars = "tau")

```

<br>

##### Eta (Case transformation)

```{r, message = F}

model_fit_summary("eta") %>% select(-true)
traceplot(obj = model_objs$model_draws, "^eta")
mcmc_hist(model_objs$model_draws, pars = "eta")

```

<br>

##### Epsilon (case underreporting)

```{r, message = F}

purrr::map(str_c("e_i[", 1:M, "]"), model_fit_summary) |> 
  set_names(str_c("e_i[", 1:M, "]")) |> 
  bind_rows(.id = "Parameter") |> select(-true) |> 
  print(n=M)

```

<br>

##### Delta (persistence)

```{r, message = F}

model_fit_summary("delta") %>% select(-true)
traceplot(obj = model_objs$model_draws, "delta")
mcmc_hist(model_objs$model_draws, pars = "delta")

```

<br>

##### Omega (spatial weights):

<br>

**Rhat**

<br>

```{r, message = F, fig.height = 10}

w_pars <- model_objs$metadata$model_params[str_which(model_objs$metadata$model_params, "^w\\[")]

w_rhats <- model_objs$rhats[w_pars]
mcmc_rhat(w_rhats) + 
  theme_classic() + 
  theme(axis.text.y = element_blank(), 
        axis.ticks = element_blank())

```

<br>

```{r}
if(length(names(which(w_rhats >=1.1)))>0){
  print(c("Rhat >= 1.1:", names(purrr::map(which(w_rhats >=1.1), ~{
    str_c("edge ", .x, ": ", 
          str_c(countries_subset[which(spatial_objs$map_from_edge==.x, arr.ind = T)], 
                collapse = ", "))
  }) |> set_names())))
}else{
  print("All Rhats <1.1")
}
```

<br>

**Neff**

<br>

```{r, message = F, fig.height = 10}

w_neffs <- model_objs$neffs[w_pars]
mcmc_neff(w_neffs) + 
  theme_classic() + 
  theme(axis.text.y = element_blank(), 
        axis.ticks = element_blank())

```

<br>

```{r}
if(length(names(which(w_neffs <=0.1)))>0){
  print(c("Neff ratio <= 0.1:", names(purrr::map(which(w_neffs <=0.1), ~{
    str_c("edge ", .x, ": ", 
          str_c(countries_subset[which(spatial_objs$map_from_edge==.x, arr.ind = T)], 
                collapse = ", "))
  }) |> set_names())))
}else{
  print("All Neff Ratios >0.1")
}
```

<br>

#### Initial probabilities of presence

<br>

**Rhats:**

<br>

```{r}

pi_pars <- model_objs$metadata$model_params[str_which(model_objs$metadata$model_params, "^pi_0\\[")]

pi_rhat <- model_objs$rhats[pi_pars]
mcmc_rhat(pi_rhat) + 
  theme_classic() + 
  theme(axis.text.y = element_blank(), 
        axis.ticks = element_blank())

```

<br>

**Neffs:**

<br>

```{r}

pi_neff <- model_objs$neffs[pi_pars]
mcmc_neff(pi_neff) + 
  theme_classic() + 
  theme(axis.text.y = element_blank(), 
        axis.ticks = element_blank())

```

<br>

**Probabilities:**

<br>

```{r, fig.height = 14, fig.width = 10}

poststan_df("pi_0", state = F) |> 
  mutate(location = str_c("loc-", ind),
         time = 1) |> 
  translate_cols() |> 
  select(country, te, year, mean) |> 
  ggplot() +
  geom_point(aes(x = te, y = mean, color = te)) +
  facet_wrap(~country, ncol = 4) +
  plot_colors +
  theme_bw() +
  scale_y_continuous(limits = c(0,1))

```


<br>

#### Introduction probabilities (gamma) by country & strain over time:

<br>

```{r, fig.height = 18}

pred_gamma <- translate_cols(poststan_df("gamma", state = F))

ggplot(data = pred_gamma) +
    geom_ribbon(aes(x = year, ymin = q5, ymax = q95, fill = te), alpha = .1, position = position_dodge(0.2)) +
    geom_line(aes(x = year, y = mean, color = te), position = position_dodge(0.2)) +
    facet_wrap(~country, ncol = 2) +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    labs(color = "TE",
    fill = "TE") + 
    plot_colors +
    plot_fillcolors +
    guides(color = guide_legend(nrow = 2), fill =guide_legend(nrow = 2)) +
    theme_bw() +
    theme(text = element_text(size = 8), 
          legend.position = "bottom",
          ggside.axis.text.y = element_blank())

pred_gamma_tot <- pred_gamma |> 
  group_by(te, year) |> 
  mutate(prob = sum(mean)) |> 
  ungroup() |> 
  select(te, year, prob) |> 
  distinct() |> 
  arrange(te, year)

ggplot() +
  geom_line(aes(x = year, y = prob, color = te), 
            position = position_dodge(0.2), data = pred_gamma_tot) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  labs(color = "TE",
  fill = "TE") + 
  plot_colors +
  guides(color = guide_legend(nrow = 2), fill =guide_legend(nrow = 2)) +
  theme_bw() +
  theme(text = element_text(size = 8), 
        legend.position = "bottom",
        ggside.axis.text.y = element_blank())

```


